A Study of Common Bug Fix Patterns in Rust

### 2022-11-23 methodology.tex
added changes

% \todo{introductory para saying that we're going to describe our methodology}

In this section, we describe our methodology, providing full details in later subsections. Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs which parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints. After conducting a manual analysis on the obtained clusters, we refine them towards manifesting bug fix patterns.

We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST). A change pattern is a set of code changes that serves a specific purpose. This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns). In this work, we are interested in bug-fixing patterns. Thus, from now on, whenever we refer to code patterns, we mean bug-fixing patterns.

@@

% \todo[inline]{I think we only need to talk about this next bit in a subsection. It's an implementation detail, not fundamental to our approach: ``We decided to implement our pipeline completely in Python. To make this work, we need a way to get an AST as a Python dictionary.''}

% \todo[inline]{I wonder if this should really go here, or if there should be a shorter version.}
% To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it. After parsing these two revisions, we will have two ASTs. Tree diff algorithms can compute the differences between two arbitrary trees. When the trees are programs' abstract syntax trees, we call their diff an ASTDiff. 
% , encoded in two Python dictionaries; We use dictdiffer to obtain ASTDiffs.

We decided to implement Ruxanne completely in Python. Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021. We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:data_modelling} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth. Then, we applied the DBSCAN clustering algorithm~\citep{ester1996density} on the resulting datapoints. Section~\ref{sec:clustering_data} describes clustering in detail, justifies our choice of DBSCAN for clustering, and explains how we tuned it to improve cluster quality.

\subsection{\label{sec:data_modelling}Data modelling}

% \todo[inline]{Maybe this next paragraph is what you want at the top of 2.1. Then in the next next paragraph you put a shorter description of the pipeline, in a sentence before ``we mined'', and you reference 2.1 there. Anyway think about organization and try to have something that makes sense.}
To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it. After parsing these two revisions, we will have two ASTs. Tree diff algorithms can compute the differences between two arbitrary trees. When the trees are programs' abstract syntax trees, we call their diff an ASTDiff. An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit. Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints. Section~\ref{sec:parsing_programs} describes our program parsing process, and how we obtained ASTDiff from code changes. Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information out of an ASTDiff, which then allowed us to obtain clusters that contained similar datapoints.

% \todo[inline]{should always be text between subsection and subsubsection. This is the per-change workflow, right?}

@@

The output of dictdiffer is a list of diffs, where each diff comprises three parts. The first part specifies the modification type:

@@

The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change---terminals and nonterminals. Ruxanne looks for changes within each root-level scope (Rust book calls these root-level scopes the \textit{items} of a crate\footnote{\url{https://doc.rust-lang.org/reference/items.html}}). That is, our assumption is that each change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, our tool detects both continuous and non-continuous line changes within one scope.

Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one element contributing to the change. A Depth First Search allows us to collect all paths, and we return only the paths that we store in our datapoints (i.e. touching $\mathtt{NT} \cup \mathtt{BCE}$).
% \todo{if you mathtt BCE here, you should do it everywhere, i.e. in the paragraph at the start of 2.1.1.}

Figure~\ref{fig:extraction} illustrates path extraction on a simple change. The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. We found it crucial to record the order of nodes in the path, as this allows us to record that order in our datapoint embedding.

\paragraph{Code Embedding.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\citep{xu2005survey}. The number of columns controls the amount of information we can embed in the datapoints.

@@

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn. However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs. In Section~\ref{sec:weighting_scheme}, we provide and example on how this approach works.

@@

    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights, resulting in the essence of change (visualized with a circle pack figure at bottom right). Vectors show (above) the number of occurrences and (below) weighted occurrence counts.}
\end{figure}

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.

We made two observations about changes. First, we saw that an item (non terminal or BC-related element) that occurs closer to AST leaves tends to be more semantically important. Relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus. Specifically, we mined the last 20 bug related commits of the last 20 most-starred Rust projects on GitHub, and ran them through our pipeline. We recorded the total number of occurrences of each item. Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$. Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e. closer to the leaves. Items that occur often and have lower weights are, we believe, less important in describing a change.

@@

\begin{table}
\caption{\label{table:repos} Target repositories for mining}
\begin{tabular}{l r r r r}
    Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\
    denoland/deno & 85.9 & 146055 & 8026 & 49.22 \\
    tauri-apps/tauri & 52.5 & 50014 & 3041 & 31.67 \\
    alacritty/alacritty & 42.4 & 30124 & 2036 & 27.53 \\
    sharkdp/bat & 37.6 & 9496 & 2457 & 59.13 \\
    starship/starship & 29.3 & 34888 & 2379 & 25.1 \\
    rust-lang/rustlings & 30.6 & 5067 & 1466 & 8.92 \\
    meilisearch/meilisearch & 30.2 & 25981 & 4029 & 279.7 \\
    sharkdp/fd & 24.9 & 6542 & 1068 & 12.2 \\
    swc-project/swc & 24.2 & 346501 & 5884 & 44.34 \\
    yewstack/yew & 24.5 & 44295 & 2280 & 21.79 \\
    xi-editor/xi-editor & 19.6 & 39292 & 2105 & 28.67 \\
    AppFlowy-IO/AppFlowy & 28.4 & 57646 & 3321 & 29.49 \\
    firecracker-microvm/firecracker & 19.6 & 82021 & 3620 & 58.26 \\
    nushell/nushell & 21.1 & 163562 & 6088 & 22.38 \\
    ogham/exa & 19.5 & 10852 & 1558 & 11.99 \\
    SergioBenitez/Rocket & 18.7 & 68952 & 2135 & 19.75 \\
    rustdesk/rustdesk & 30.2 & 48115 & 2620 & 72.81 \\
    tokio-rs/tokio & 17.9 & 121400 & 3101 & 28.87 
\end{tabular}
\end{table}

@@

Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns. Like \cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons. Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm. Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters. We did try k-means and its more powerful fuzzy variant Fuzzy C-means~\citep{dunn1973fuzzy} and did not get useful clusters.
\todo[inline]{Wouldn't the reviewers complain why you haven't put k-means data in the paper?}

@@

% TODO fix this caption
\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{clusters.pdf}
\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows. In each subplot, the two blue bars in the left respectively indicate the number of clustered ($G_c$) and noise ($G_n)$ datapoints (the number is shown above each bar) after applying clustering on general database $D_g$. The two orange bars in the right show similar data but for the borrow-checker database $D_{b}$. In each subplot's legend, $C_g$ and $C_b$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
\end{figure}

@@

Since both large and small values of $\epsilon$ would steer us away from obtaining meaningful clusters, we felt the need to carry out a manual analysis of clusters for better parameter tuning. In the manual analysis, we randomly pick 50 clusters; from each cluster we randomly choose 10 datapoints. The person analyzing the cluster (The first author) then looks at the code and specifies the clusters that contained more than five datapoints showing similar patterns. After conducting the manual analysis, we decided to use $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, as our clustering parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively. As a final step, we exclude clusters that contained datapoints from fewer than 3 projects. That is because we wanted our clusters to contain cross-project (e.g. potentially generalizable) patterns as much as possible.

@@

In this work, we were only interested in the first two groups and ignored the clusters for refactoring changes. Also, we prefer reporting clusters that manifest more-specific patterns. An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs. For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Changing a statement in a function's body':

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had fewer than 50 datapoints, we would analyze all of the datapoints in the cluster. Regarding general patterns, we obtained 577 clusters, out of which 428 included cross-project patterns. Out of these 428 clusters, 35 clusters included more than 50 points. That is, we did not do random sampling for 393 clusters, we just manually analyzed all the datapoints in them. Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project pattern. Only 4 of these 49 clusters included more than 50 datapoints, therefore, we only did a random sampling for 4 clusters. For each cluster, we read the code and the bug reports of the datapoints (if they existed). If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster. Considering the descriptions, we linked similar clusters together as possible candidates for merging. We considered these candidates and carried out appropriate ones. Finally, we selected 20 of the remaining clusters---the ones that manifested more-specific changes. Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns} present our clusters.

>>>>

In this section, we describe our methodology, providing full details in later subsections. Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs. This pipeline parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints. After conducting a manual analysis on the obtained clusters, we refine them to obtain our proposed bug fix patterns.

We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST). A change pattern is a set of code changes that serves a specific purpose. Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) as well as improving code readability or maintainability (refactoring patterns). In this work, we are interested in bug-fixing patterns. Thus, from now on, whenever we refer to code patterns, we mean bug-fixing patterns.

@@

We implemented Ruxanne completely in Python. Our target repositories were the top 18 most-starred open source Rust projects on GitHub as of August 2021. We mined (using Pydriller) all the bug related commits and ran them through our pipeline; Section~\ref{sec:code_embedding} describes how we extracted fixed-size datapoints from the code changes, and Section~\ref{sec:mining} describes our mining methodology in depth. Then, we applied the DBSCAN clustering algorithm~\citep{ester1996density} on the resulting datapoints. Section~\ref{sec:clustering_data} describes clustering in detail, justifies our choice of DBSCAN for clustering, and explains how we tuned it to improve cluster quality.

\subsection{\label{sec:code_embedding}Code Embedding}

To compute the contents of a change, we need two code revisions: the revision before the change, and the revision after it. After parsing these two revisions, we will have two ASTs. Tree diff algorithms can compute the differences between two arbitrary trees. When the trees are programs' abstract syntax trees, we call their diff an ASTDiff. An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit. Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints. Section~\ref{sec:parsing_programs} describes our program parsing process, and how we obtained ASTDiff from code changes. Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information out of an ASTDiff, which then allowed us to obtain clusters that contained similar datapoints.

@@

The output of dictdiffer is a list of diffs, where each diff has three parts. The first part specifies the modification type:

@@

The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change---terminals and nonterminals. Ruxanne looks for changes within each root-level scope (Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate). That is, our assumption is that each change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, our tool detects both continuous and non-continuous line changes within one scope.

Looking at the tree encoded in the content part of each diff, we realized that each path to a leaf represents one element contributing to the change. A Depth First Search allows us to collect all paths, and we return only the paths that we store in our datapoints (i.e. touching $\mathtt{NT} \cup \mathtt{BCE}$).

Figure~\ref{fig:extraction} illustrates path extraction on a simple change. The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of head to the callee instead of sending head itself. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. We found it crucial to record the order of nodes in the path, so that we could record that order in our datapoint embedding.

\paragraph{Code Embedding.} A key question is how to transform these paths to a fixed size datapoint. We need fixed-size datapoints, as they are a requirement for well-known centroid-based and density-based clustering algorithms~\citep{xu2005survey}. First, we need to define a fixed set of columns. The number of columns controls the amount of information we can embed in the datapoints.

@@

We propose a novel representation. Similar to the naive approach, we order-insensitively define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ we collected from Syn. However, to simulate order-sensitiveness, we add two more steps. First, we record the number of occurrences of items within the paths at each scope. The reason behind this decision is because a different order of program elements yields a different number of occurrences of underlying dimension. Second, we semi-automatically design a weighting scheme to prioritize non terminals that we think are more salient for Rust bugs. Section~\ref{sec:weighting_scheme}, immediately below, illustrates this approach on an example..

@@

    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right. Vectors show (above) the number of occurrences and (below) weighted occurrence counts.}
\end{figure}

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.
We made two observations about changes. First, we saw that items (non terminals or BC-related elements) that occurs closer to AST leaves tend to be more semantically important. Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus. Specifically, we mined 20 of the most recent bug related commits of the top 20 most-starred Rust projects on GitHub, and ran them through our pipeline. We recorded the total number of occurrences of each item. Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$. Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e. closer to the leaves. Items that occur often and have lower weights are, we believe, less important in describing a change.

@@

\begin{table}
\caption{\label{table:repos} We chose the 18 most-starred Rust projects on GitHub as target repositories for mining.
}
\begin{tabular}{l r r r r}
    Project Name & Stars(k) & LOC & \# Commits & Avg. Code Churn \\
    denoland/deno & 85.9 & 146055 & 8026 & 49.22 \\
    tauri-apps/tauri & 52.5 & 50014 & 3041 & 31.67 \\
    alacritty/alacritty & 42.4 & 30124 & 2036 & 27.53 \\
    sharkdp/bat & 37.6 & 9496 & 2457 & 59.13 \\
    starship/starship & 29.3 & 34888 & 2379 & 25.1 \\
    rust-lang/rustlings & 30.6 & 5067 & 1466 & 8.92 \\
    meilisearch/meilisearch & 30.2 & 25981 & 4029 & 279.7 \\
    sharkdp/fd & 24.9 & 6542 & 1068 & 12.2 \\
    swc-project/swc & 24.2 & 346501 & 5884 & 44.34 \\
    yewstack/yew & 24.5 & 44295 & 2280 & 21.79 \\
    xi-editor/xi-editor & 19.6 & 39292 & 2105 & 28.67 \\
    AppFlowy-IO/AppFlowy & 28.4 & 57646 & 3321 & 29.49 \\
    firecracker-microvm/firecracker & 19.6 & 82021 & 3620 & 58.26 \\
    nushell/nushell & 21.1 & 163562 & 6088 & 22.38 \\
    ogham/exa & 19.5 & 10852 & 1558 & 11.99 \\
    SergioBenitez/Rocket & 18.7 & 68952 & 2135 & 19.75 \\
    rustdesk/rustdesk & 30.2 & 48115 & 2620 & 72.81 \\
    tokio-rs/tokio & 17.9 & 121400 & 3101 & 28.87 
\end{tabular}
\end{table}

@@

Having collected our datapoints, our next task is to cluster them so that we can categorize bug patterns. Like \cite{hanam2016discovering}, we use the DBSCAN clustering algorithm for two main reasons. Firstly, in contrast to k-means, it does not require the number of clusters in advance as an input to the algorithm. Secondly, it is a density-based clustering method, which means that it detects arbitrarily shaped clusters; centroid-based methods like k-means can't detect such clusters.

@@

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{clusters.pdf}
\caption{\label{fig:clustering} Clustering results with nine different combinations of parameters $\epsilon$ and $Z$, respectively separated in columns and rows. In each subplot, the two blue bars in the left respectively indicate the number of clustered ($G_c$) and noise ($G_n)$ datapoints (the number is shown above each bar) after applying clustering on general database $D_g$. The two orange bars in the right show similar data but for the borrow-checker database $D_{b}$. In each subplot's legend, $C_g$ and $C_b$ show the number of clusters obtained from $D_g$ and $D_b$, respectively}
\end{figure}

@@

Since both large and small values of $\epsilon$ would steer us away from obtaining meaningful clusters, we felt the need to carry out a manual analysis of clusters for better parameter tuning. In the manual analysis, we randomly picked 50 clusters; from each cluster we randomly chose 10 datapoints. The person analyzing the cluster (the first author) then looked at the code and specified the clusters that contained more than five datapoints showing similar patterns. After conducting the manual analysis, we decided to use $\epsilon=0.0001, Z=5$ and $\epsilon=0.001, Z=5$, as our clustering parameters for $D_g$ and $D_b$, resulting in 577 and 102 clusters, respectively. As a final step, we excluded clusters that contained datapoints from fewer than 3 projects. That is because we wanted our clusters to contain cross-project (e.g. potentially generalizable) patterns as much as possible.

@@

In this work, we were only interested in the first two groups and ignored clusters for refactoring changes. Also, we prefer reporting clusters that manifest more-specific patterns. An abstract pattern as opposed to a more-specific pattern can account for many changes in the programs. For instance, we deem the pattern `Changing a clone of a variable to a borrowing of it in a function argument' more specific than `Changing a statement in a function's body':

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had fewer than 50 datapoints, we analyzed all of the datapoints in the cluster. Regarding general patterns, we obtained 577 clusters, out of which 428 included cross-project patterns. Of these 428 clusters, 35 clusters included more than 50 points. That is, we did not do random sampling for 393 clusters, we just manually analyzed all the datapoints in them. Regarding BC-related patterns, out of 102 obtained clusters, 49 clusters included cross-project patterns. Only 4 of these 49 clusters included more than 50 datapoints, therefore, we only did a random sampling for 4 clusters. For each cluster, we read the code and the bug reports of the datapoints (if they existed). If we judged that the members of a cluster fit into a pattern, we wrote a natural language description of that cluster. Considering the descriptions, we linked similar clusters together as possible candidates for merging. We considered these candidates and carried out appropriate ones. Finally, we selected 20 of the remaining clusters---the ones that manifested more-specific changes. Sections~\ref{sec:common_patterns}~and~\ref{sec:bc_patterns} present our clusters.

### 2022-11-23 results.tex
added changes

In this section, we present the clusters of fix patterns that we found in our data set. \RQG{} is about the borrow-checker related patterns, while \RQBC{} is about general patterns. However, our technique fundamentally relies on our data modelling approach successfully capturing the most important aspects of program changes. We thus first evaluate the data modelling approach as our RQ1, before continuing with \RQBC{} and \RQG{}.

\subsection{Evaluating the data modelling approach (RQ1)}

\textbf{Does our data modelling approach capture the most important aspect of program changes?}

As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (a weighting scheme) while creating our datapoints to make the most important program elements stand out from all the non terminals seen in the AST (along with some BC-related elements). Here we evaluate the effectiveness of our code embedding.

The general idea for our evaluation is determining whether the most important program elements get high values in our embedding, i.e. are recognized as important. The analyzer compares the actual change in the code and compares it to the visual representation of the respective data point. We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle of each non terminal shows its importance in our embedding (as seen in Figure~\ref{fig:essence}).

@@

We carried out a manual experiment for evaluating our code embedding. In the experiment, the analyzer should rank the most important element based on the provided circlepack figure and the change's source code. Based on these ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$. Both authors carried out the experiment independently. We report the results as the portion of the changes which both authors agreed upon its embedding correctness. The final results are shown in Figure~\ref{fig:rq1}.

@@

Although, on average, reporting most important element within the top four elements in 90\% of the cases is an accpectable outcome, we examined our pipeline for the other 10\%. An analysis shows that our pipeline fails in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). It is worth mentioning that these changes tend to be less cross-project as having many changed elements make it easier for datapoints to be dissimlar.

@@

We collected 60583 datapoints for general fix patterns using a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to compute both $D_g$ and $D_b$. We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters. After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}. As it is shown in the table, we have separated the patterns based on the underlying program element (Category column). We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column). Now we provide detailed description for each of these patterns:

@@

\noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules. In the instances of this cluster, adding \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access. On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g. for encapsulation purposes.

@@

We collected 27143 datapoints for borrow-checker related fix patterns. We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters. After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}. Here we provide detailed description for each of these patterns:

>>>>

In this section, we present the clusters of fix patterns that we found in our data set. RQ2 is about general patterns, while RQ3 is about borrow-checker related patterns. However, our technique fundamentally relies on our code embedding approach successfully capturing the most important aspects of program changes. We thus first evaluate the code embedding approach as our RQ1, before continuing with RQ2 and RQ3.

\subsection{Evaluating the code embedding approach (RQ1)}

\textbf{Does our code embedding approach capture the most important aspect of program changes?}

As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST. In addition to that, we also incorporated the BC-related elements in our datapoints. Here we evaluate the effectiveness of our embedding.

Our evaluation aims to determine whether the most important program elements get high values in our embedding, i.e. are recognized as important. A human analyzer compares the actual change in the code and compares it to the visual representation of the respective data point. We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle for each non terminal shows its importance in our embedding (Figure~\ref{fig:essence} presented an example of the circle pack figures).

@@

The human looks at the program output and assigns a rank to it

We carried out a manual experiment for evaluating our code embedding. In the experiment, the human analyzer looks at the provided circlepack figure and the change's source code and assigns a rank to it. This rank reflects the rank of the most important element in the circle pack figure according to the analyzer's view. Based on these ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$. Both authors acted as analyzer and carried out the experiment independently. We report the results as the fraction of the changes on which both authors agreed upon its embedding correctness. The final results are shown in Figure~\ref{fig:rq1}.

@@

Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 90\% of the time---we investigated the other 10\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.

@@

We collected 60583 datapoints for general fix patterns using a server with a Xeon Gold 5120 processor (Q3-2017, 14-core 2.2 to 3.2GHz) and the data collection took around 26 hours to compute both $D_g$ and $D_b$. We used $\epsilon=0.0001$ and $Z=5$ as our DBSCAN parameters and that resulted in 577 clusters. After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 12 general patterns, which we describe in Table~\ref{table:general}. As shown in the table, we have separated the patterns based on the underlying program element (Category column). We have given each pattern a specific ID (ID column) and determined the number of instances manifesting that fix pattern (\# Instances column), as well as the number of projects in which the pattern was seen (\# Project column). Now we provide a detailed description for each of these patterns:

@@

\noindent\textbf{Description:} Adding \texttt{pub} to a struct field makes it possible to access that field from external modules. In the instances of this cluster, adding the \texttt{pub} access modifier happened in bug fixing changes, where the developer needed to access a field which had not been marked for public access. On the other hand, the public access might be revoked if the developer realizes there are no external accesses to the field, and that such accesses are not desirable, e.g. for encapsulation purposes.

@@

We collected 27143 datapoints for borrow-checker related fix patterns. We used $\epsilon=0.001$ and $Z=5$ as our DBSCAN parameters and that resulted in 102 clusters. After manual analysis, as described in~\ref{sec:manual_analysis_cluster_selection}, we ended up with 8 borrow-checker related fix patterns, which we, similar to the previous table, describe in Table~\ref{table:bc}. Here we provide a detailed description for each of these patterns:
